Type Refinements for the Working Class

### 2015-11-09 paper.tex
begin developing meaning explanation for refinement

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsWf[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
}

\newcommand\IsBTm[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\IMode{#5}%
}

@@

@@

>>>>

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsBTm[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}:\IMode{#5}%
}

@@

\newcommand\Refines[3]{\IMode{#1}\parallel \IMode{#2} \sqsubset \OMode{#3}}
\newcommand\GRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}
\newcommand\RefinesCtx[2]{\IMode{#1}\sqsubset^\star\OMode{#2}}

@@

\section{Abstract Binding Trees and Symbols}

See \cite{sterling-morrison:2015} for the development of abstract binding trees
with symbols.

TODO: give a brief description of the framework, and present its rules.

\section{Behavioral Refinements}

Fixing a signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ in the abt
framework, we will define the notion of behavioral refinement by propounding
several judgments and their semantical explanations. We will require two
judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$, which means that
$\Phi$ is a context which refines the symbol context $\Upsilon$ (presupposing
$\IsSymCtx{\Upsilon}$, and $\Refines{\Phi}{\phi}{\tau}$, which shall presuppose
$\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).

\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathsf{nil}}}
\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathsf{snoc}}}

The first judgment we define inductively:
\begin{gather*}
  \infer[\RefinesCtxNil]{
    \RefinesCtx{\cdot}{\cdot}
  }{
  }\qquad
  \infer[\RefinesCtxSnoc]{
    \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
  }{
    \RefinesCtx{\Phi}{\Upsilon} &
    \Refines{\Phi}{\phi}{\tau}
  }
\end{gather*}

The second judgment is defined coinductively via the following meaning
explanation:
\begin{quote}
  To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
  $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
  $\Of{\varrho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, for any $M, N$,
  what it means for $M$ and $N$ to be identified under $\phi$, supposing
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
  $\IMode{M}\sim_\phi\IMode{N}$ then $\IMode{N}\sim_\phi\IMode{M}$, and if
  $\IMode{M}\sim_\phi\IMode{N}$ and $\IMode{N}\sim_\phi\IMode{O}$, then
  $\IMode{M}\sim_\phi\IMode{O}$.
\end{quote}

In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
$\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
$\Upsilon'$, which partial equivalence relation on the closed $\Upsilon'$-terms
of $\tau$ that $\phi$ denotes. At this point, it should be remarked that this
is very similar to the semantical explanation of typehood given in
\cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
setting, and that we have fibred the entire apparatus over collections
$\Upsilon$ of symbols.

\begin{thm}[Weakening]
  If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
  $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
\end{thm}
\begin{proof}
  This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
  $\Upsilon\hookrightarrow\Upsilon,u:\sigma$.
\end{proof}

When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ is an
\emph{global} refinement and may simply write $\GRefines{\phi}{\tau}$. By
weakening, if $\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
$\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.
